---
title: "Homework 2"
author: "Cong Zhang"
date: 2020-09-29
output: github_document
---

This is my solution to Homework 2.

```{r message = FALSE}
library(tidyverse)
library(readxl)
```


## Problem 1

First, define a path to the dataset. 

```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```


Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

Read precipitation data! For 2018 and 2017. 

```{r}
precip_2018 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
precip_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation dataframes. In the following code chunk, I create a "helper" tibble that contains pairs of numeric and character ways of representing month, and then merge that (using month number as a key) with the precipitation dataset. This technique is one I use often when I need to recode a moderate or large number of values for a variable. 

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)
precip_df =
	left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


## Problem 2

Read the NYC Transit data. 

```{r nyctransit_df, message = FALSE}
nyctransit_df = 
	read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
	janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
	mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

The resulting NYC Transit dataset contains the following variables: `r names(nyctransit_df)`.

My data cleaning steps so far are as follows: First, import the NYC Transit data from a csv file by using the `read_csv` function. Second, clean the variable names with the `clean_names` function. Third, retain the required variables with the `select` function. Finally, change the entry variable from character to logical variable by using the `mutate` and `recode` functions.

The resulting dataset has `r nrow(nyctransit_df)` rows and `r ncol(nyctransit_df)` columns. I think the data are tidy because columns are variables; rows are observations; and every value has a cell.


Answers to the three questions:

* There are `r distinct(nyctransit_df, station_name, line) %>% nrow()` distinct stations.

* `r filter(nyctransit_df, ada == TRUE) %>% distinct(station_name, line) %>% nrow()` distinct stations are ADA compliant.

* `r filter(nyctransit_df, vending == "NO" & entry == TRUE) %>% nrow() / filter(nyctransit_df, vending == "NO") %>% nrow() * 100`% of station entrances / exits without vending allow entrance.
